---
title: "HGUTheme"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HGUTheme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(HGUTheme)
library(tidyverse)
```


`HGUTheme` has three main classes of functions:

* Functions to return the HEX code of the colours used in the unit logos
* Functions to create palettes interpolating the colours used in the unit logos
* Functions to scale `ggplot2` plot colour and fill aesthetics to use interpolated colour palettes from the unit logos

These are explained below with examples.

## Returning Colour HEX Codes

The most basic functions return the colour HEX codes. This can be seen here

```{r return UKRI light blue}
ukri_light_blue()
```

This is useful to set colours for various aspects of plots or visualisations

```{r}
barplot(
  3:1,
  col = ukri_light_blue()
)
```

As well as returning individual colours, palettes based on the logo can also be returned that will contain every colour in the logo

```{r}
ukri_blues()
```

This can be used to create plots such as this

```{r}
barplot(
  3:1,
  col = ukri_blues()
)
```

## Interpolating Colour Palettes

The UKRI blues palette only has three colours in it, however, what if you want to use more colours in your plot? You can also interpolate colours based on the theme, so say you want to use 5 colours from the UKRI blues you can use

```{r}
ukri_blues_n(5)
```

This can be used in plots like the following

```{r}
barplot(
  5:1,
  col = ukri_blues_n(5)
)
```

## Automatically Colour `ggplot2` Plots

`ggplot2` is a popular plotting library from the `tidyverse`. It is used to create high quality visualisations. An example of a basic `ggplot2` plot is given below

```{r}
plot_data <- tibble(
  x = LETTERS[1:5],
  y = 5:1
)

plot_data %>%
  ggplot() +
  aes(x = x, y = y, fill = x) +
  geom_col() + theme_classic() +
  guides(fill = "none") + 
  labs(x = "Group", y = "Value")
```

The plots made by `ggplot2` are created in layers. Another layer can be added from `HGUTheme` to recolour this plot with unit logo colour themes.

```{r}
plot_data %>%
  ggplot() +
  aes(x = x, y = y, fill = x) +
  geom_col() + theme_classic() +
  guides(fill = "none") + 
  labs(x = "Group", y = "Value") +
  scale_fill_ukri_d()
```

There are layers for changing the colour and the fill aesthetics of your plot. These functions end in either a *c* or a *d* indicating whether you are using a continuous or a discrete scale.

For continuous scales, v0.1.0 of `HGUTheme` will only automatically interpolate between the first and last colour of the logo, as shown here with this CRUK logo colour theme missing the light blue.

```{r}
tibble(
  x = runif(1500),
  y = runif(1500)
) %>%
  filter(
    (x - 0.5)^2 + (y - 0.5)^2 < 0.25
  ) %>%
  mutate(
    col = (x - 0.5)^2 + (y - 0.5)^2
  ) %>%
  ggplot() +
  aes(x = x, y = y, colour = col) +
  geom_point() + theme_classic() +
  guides(colour = "none") + coord_equal() +
  scale_colour_cruk_c()
```

However, adding a `midpoint` argument will trigger the full theme, this is a bug that will be ironed out in a future release.

```{r}
tibble(
  x = runif(1500),
  y = runif(1500)
) %>%
  filter(
    (x - 0.5)^2 + (y - 0.5)^2 < 0.25
  ) %>%
  mutate(
    col = (x - 0.5)^2 + (y - 0.5)^2
  ) %>%
  ggplot() +
  aes(x = x, y = y, colour = col) +
  geom_point() + theme_classic() +
  guides(colour = "none") + coord_equal() +
  scale_colour_cruk_c(midpoint = 0.1)
```

It is important to note, that these layers do not have to be used and the HEX code functions can be used to supply arguments to pre existing `ggplot2` layers for similar effects, for example if the user wanted to recreate the above plot but without the dark blue, this can be acheived via

```{r}
tibble(
  x = runif(1500),
  y = runif(1500)
) %>%
  filter(
    (x - 0.5)^2 + (y - 0.5)^2 < 0.25
  ) %>%
  mutate(
    col = (x - 0.5)^2 + (y - 0.5)^2
  ) %>%
  ggplot() +
  aes(x = x, y = y, colour = col) +
  geom_point() + theme_classic() +
  guides(colour = "none") + coord_equal() +
  scale_colour_gradient(
    low = cruk_light_blue(),
    high = cruk_pink()
  )
```
